---
title: "Animal crossing gift recommender"
author: "vina"
date: "24/03/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(patchwork)
library(tidyverse)
items <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-05/items.csv')
villagers <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-05/villagers.csv')
library(imager)
library(broman)#for converting hex colour format to numerical

```

Inspect dataset
```{r}
glimpse(villagers)
head(items)
```

Loading img URL 
```{r}

plot_villager<-function(villager_name){
  ind=which(villagers$name==villager_name)
  villager_of_choice_img <- load.image(villagers$url[ind])%>%as.data.frame(wide="c") %>% mutate(rgb.val=rgb(c.1,c.2,c.3))

plot=ggplot()  + geom_raster(data=villager_of_choice_img, aes(x=x, y=y, fill=rgb.val)) +scale_fill_identity()+scale_y_reverse()
return(plot)
}

plot_villager("Egbert")

plot_villager("Sherb")


```

Extract hex codes from center of image. This is because if we just look at which colors are available, we might end up detecting the black background the most etc so we we should focus our efforts around the centre of the villager iamge. 

```{r}
choose_villager<- function(villager_name){
  
  ind=which(villagers$name==villager_name)
villager_of_choice_img <- load.image(villagers$url[ind])%>%as.data.frame(wide="c") %>% mutate(rgb.val=rgb(c.1,c.2,c.3))

#get center coordinates 
clothes_x= round(max(villager_of_choice_img$x)/2 )
clothes_y= round(max(villager_of_choice_img$y)/2 )

# iteratively find coords and colours close to center  

up=list()
low=list()
for (i in 1:20) {
    upper=villager_of_choice_img%>%filter(x==clothes_x+i,y==clothes_y+i)
    lower=villager_of_choice_img%>%filter(x==clothes_x-i,y==clothes_y-i)
    up[[length(up) + 1]] <- upper$rgb.val
    low[[length(low) + 1]] <- lower$rgb.val
}

total_col= as_tibble(c(unlist(up),unlist(low)))

return(total_col)
}


print(choose_villager("Egbert"))


```


Now over to items. 


```{r}
#omit no source info as we want to know how to get desired items

gift_type<- function(gift_type){
  choice=items%>%filter(category==gift_type)
  return(choice)
}

gift_choice= gift_type("Tops")

#loop through items and extract colors- this is the same procedure as what we did for villagers earlier. The only differece here is that we need a double loop that will first loop through each desired  gift item and then over corrdinates/colours to extract centre colours. 

tot=list()
for (x in 1:length(gift_choice)) {
    img <- load.image(gift_choice$image_url[x])%>%as.data.frame(wide="c") %>% mutate(rgb.val=rgb(c.1,c.2,c.3))
    clothes_x= round(max(img$x)/2 )
    clothes_y= round(max(img$y)/2 )
    up=list()
    low=list()
    for (i in 1:20) {
        upper=img%>%filter(x==clothes_x+i,y==clothes_y+i)
        lower=img%>%filter(x==clothes_x-i,y==clothes_y-i)
        up[[length(up) + 1]] <- upper$rgb.val
        low[[length(low) + 1]] <- lower$rgb.val
    }
    
    total_col= as_tibble(c(unlist(up),unlist(low)))
    tot[[x]]=total_col
    colnames(tot[[x]])<-gift_choice$name[x]
    
}

tot=flatten_dfc(tot)

head(tot)
```

convert hex to decimal values so that we can quantitatively comapre colours between villagers and items to find nearest matches. 

```{r}

#convert hex to dec for items of choice :
for (x in 1:length(tot)) {
    tot[x][[1]]=gsub("#","",tot[x][[1]])
    tot[x][[1]]= broman::hex2dec(tot[x][[1]])
}

tot=pivot_longer(tot, everything())

#convert hex to dec for villager of choice:
fav_villager=choose_villager("Egbert")

median_col=list()
for (x in 1:nrow(fav_villager)) {
    y=gsub("#","",fav_villager$value[x])
    col_convert=broman::hex2dec(y)
   median_col[x]=col_convert
    
}
```


Matching villager and item colors. This was a difficult process because colors were so varied and sometimes unique that they did not automatically match. As a compromise I decided to just find the median color value as this was most representative of those center colors. After computing median colour value, I looked through the items colours to find the closest match and this was the recommended gift item. 

```{r}

median= median_col%>%unlist()%>%as_tibble()%>%summarise(median(value))

closest<-function(xv,sv){
    xv[which(abs(xv-sv)==min(abs(xv-sv)))] }


tot$name[which.min(abs(tot$value- median$`median(value)`))]


```

main play- putting all relevant previous function together into a large function. This main play function does not only output the name of chosen gift but also prints a ggplot of the gift+additional details like price. 

```{r}

main_play<-function(gift,villager){
    gift_choice= gift_type(gift)
    
    #loop through items and extract colors  
    tot=list()
    for (x in 1:length(gift_choice)) {
        
        img <- load.image(gift_choice$image_url[x])%>%as.data.frame(wide="c") %>% mutate(rgb.val=rgb(c.1,c.2,c.3))
        clothes_x= round(max(img$x)/2 )
        clothes_y= round(max(img$y)/2 )
        up=list()
        low=list()
        for (i in 1:20) {
            upper=img%>%filter(x==clothes_x+i,y==clothes_y+i)
            lower=img%>%filter(x==clothes_x-i,y==clothes_y-i)
            up[[length(up) + 1]] <- upper$rgb.val
            low[[length(low) + 1]] <- lower$rgb.val
        }
        
        total_col= as_tibble(c(unlist(up),unlist(low)))
        tot[[x]]=total_col
        colnames(tot[[x]])<-gift_choice$name[x]
        
    }
    
    tot=flatten_dfc(tot)
    
    #convert to decimal for item colours
    
    for (x in 1:length(tot)) {
        tot[x][[1]]=gsub("#","",tot[x][[1]])
        tot[x][[1]]= broman::hex2dec(tot[x][[1]])
    }
    
    tot=pivot_longer(tot, everything())
    fav_villager=choose_villager(villager)
    
    median_col=list()
    for (x in 1:nrow(fav_villager)) {
        y=gsub("#","",fav_villager$value[x])
        col_convert=broman::hex2dec(y)
        median_col[x]=col_convert
        
    }
    
    
    median= median_col%>%unlist()%>%as_tibble()%>%summarise(median(value))
    
    closest<-function(xv,sv){
        xv[which(abs(xv-sv)==min(abs(xv-sv)))] }
    
    
    out=as.character(tot$name[which.min(abs(tot$value- median$`median(value)`))])
    
    plot1=items%>%filter(name==out)
    p1=load.image(plot1$image_url)%>%as.data.frame(wide="c") %>% mutate(rgb.val=rgb(c.1,c.2,c.3))%>% ggplot()+ geom_raster(aes(x=x, y=y, fill=rgb.val)) +scale_fill_identity()+scale_y_reverse()+theme_void()+ggtitle(paste0(plot1$name,"/", plot1$sources, "/", plot1$buy_value," ","bells"))+ theme(plot.title = element_text(size = 8, face = "bold"))
    return(p1)
    
}





```


Trying it out
```{r}
p1=main_play("Tops","Egbert")
p2=main_play("Hats","Egbert")
p3=main_play("Fossils","Egbert")
p4=main_play("Accessories","Egbert")
p5=main_play("Bottoms","Egbert")
p6=main_play("Dresses","Egbert")

p1+p2+p3

p4+p5+p6
```

```{r}
S1=main_play("Tops","Sherb")
S2=main_play("Hats","Sherb")
S3=main_play("Fossils","Sherb")
S4=main_play("Accessories","Sherb")
S5=main_play("Bottoms","Sherb")
S6=main_play("Dresses","Sherb")

S1+S2+S3

S4+S5+S6
```
