---
title: "Movie script analyses"
author: "Vina"
date: "28/01/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(tidytext) 
library(textdata)
library(wordcloud)
library(patchwork)
library(ggpomological)
library(reactable)
```

Read dataset in and get rid of missing lines.
```{r,include=TRUE}
her <- read_excel("her.xlsx")%>%na.omit()
```

Now it is crucial to clean this dataset because it is not very informative as is. I envisioned that my ideal dataset would have a column for dialogue, speaker and scene. The first step was therefore to identify useful patterns pertaining the following variables. 

```{r}
reactable(her)
```

One easily identifiable pattern in the script is that the speaker and scene info are in all caps followed by corresponding dialogue. So the first thing to do is to extract positions within the dataset where  these all caps strings occur and then bind the dialogue between them. 

```{r}
cap=sapply(stringr::str_extract_all(her$Her, '[A-Z]{2,}'),paste0, collapse = ' ')
#extract all caps
print(cap[1:20]) #example
int=which(cap!="")#identify positions where a string was extracted
print(int[1:20])#example
```

Okay lets begin by concatenating text between each all scene/speaker. 

```{r}
i<-0
dialogoue<-c()#define list

for (x in 1:length(int)) {
  i<-i+1
  first=as.integer(int[i]+1)
  second=as.integer(int[i+1]-1)
  if (is.na(second)==FALSE) {
    t=paste0(her$Her[first:second],collapse = " ")
    dialogoue[i]=t #add to list
    
  }
}

#stage directions are usually embedded within brackets inside dialogue so we can remove that like below
dialogoue<- gsub("\\s*\\([^\\)]+\\)","",as.character(dialogoue))

dialogoue[1:20]#example

```

Nice! Now we have chunks of dialogue. Let's try to identify the speaker. Well we don't have to search very far! This is because we did extract the speaker in an earlier step where we identified all caps strings with a similar method from above. 

```{r}
speaker<-c()
i<-0
for (x in int){
  i<-i+1
  speaker[i]=cap[x]
  
}
speaker<-speaker[1:length(speaker)-1]
her_script=cbind(dialogoue,speaker)%>%as_tibble()
reactable(her_script)#output
```
We're getting somewhere good at this point but recall how we also wanted another column with details on scene/locations
```{r}
her_script<-her_script%>% mutate(scene_pos=case_when(str_detect(her_script$speaker,"INT")|str_starts(her_script$speaker, "EXT", negate = FALSE)==TRUE~her_script$speaker)) #  only if INT or EXT detected then extract scene info from speaker column and mutate to scene_pos column
reactable(her_script)
```

```{r}
#use below to fill in NAs with previous scene in list:
is = seq_along(her_script$scene_pos)
i = cummax((!is.na(her_script$scene_pos)) * is) #identify whether a scene is present or not (1 for present, 0 for NA) and multiply by is to get index value and use cummax to return the maximum of the vector from the beginning of the vector up to the current index
her_script$scene_pos[i][1:20]

```
attach scene info from above to dataset and perform additional cleaning 
```{r}
her_script<-her_script%>%filter(row_number()>1)%>%mutate(scene=paste(her_script$scene_pos[i],i[2:length(i)])) %>% mutate(index=i[2:length(i)])#remove first row and assign to scene + index so we can keep order of scenes for plotting purposes later

#remove rows that have all caps strings because these usually mean stage directions 
her_script=her_script %>% 
  filter(!str_detect(dialogoue, '[A-Z]{2,}'))%>%select(!scene_pos)
#output
reactable(her_script)
```

It is difficult to completely remove stage directions because they have no special patterns so this is a work in progress!

Let's have some fun now: its plotting time!

```{r}
#plotting longest  scenes
most_scene<-her_script%>% group_by(scene)%>%count(scene)%>%arrange(desc(n))
most_scene[1:5,]%>%ggplot(aes(scene,n,fill=scene))+geom_col()+theme_pomological()+
ggpomological::scale_fill_pomological()+theme(axis.title.x=element_blank(), axis.title.y=element_blank(),axis.text.x=element_blank(), axis.ticks.x=element_blank())
```

Most of the movie seems to take place in Theodore's bedroom which is pretty interesting because that is where he converses most with his AI partner, Samantha. We see also that most scenes are indoors either in a home or office setting. 

It will be fun to also plot a timeline of scenes so we can observe how locations/time of day changes throughout the movie.

Set parameters 
```{r}
scene_timeline<- cbind.data.frame(scene=unique(her_script$scene), scene_num=1:length((unique(her_script$scene)))) #add consistent index to determine spacing on timeline 


positions <- c(0.5, -0.5, 1.0, -1.0) #length of geom-segemnt line 
directions <- c(1, -1) #above or below y axis

line_pos <- data.frame(
  "scene_num"=unique(scene_timeline$scene_num),
  "position"=rep(positions, length.out=length(unique(scene_timeline$scene_num))),
  "direction"=rep(directions, length.out=length(unique(scene_timeline$scene_num)))
)

df <- full_join(x=scene_timeline, y=line_pos, by="scene_num", all = TRUE)


```

plotting timeline of scenes
```{r}
#define timeline plot function:
plot_timeline<-function(x_timeline,y_timeline){
  plt=ggplot(df[x_timeline:y_timeline,],aes(x=seq_along(df[x_timeline:y_timeline,]$scene_num),y=0, col=as.factor(scene_num), label="h"))+geom_hline(yintercept=0, color = "black", size=0.3)+geom_point(aes(scene_num-(x_timeline-1),0),size=6,alpha=0.5)+geom_segment(aes(y=position,yend=0,xend=scene_num-(x_timeline-1)),color='black', size=0.5)+geom_text(aes(y=position,label=scene),size=2.5)+theme_pomological()+theme(legend.position = "none")+geom_text(aes(x=scene_num-(x_timeline-1),y=-0.1,label=paste("scene",scene_num)),color="black",size=2.5) + scale_colour_pomological()+theme(axis.title.x=element_blank(), axis.title.y=element_blank(),axis.text.x=element_blank(), axis.ticks.x=element_blank(),axis.text.y=element_blank())
  return(plt)}
#plot first 8 scenes on a timeline
plot_timeline(1,8)

plot_timeline(80,87)#plot last few scenes
```

My goal now is to make sentiment plots so we can track words related to joy or sadness as the plot advances. 

```{r}
sam_dialogue<- her_script%>%filter(speaker=="SAMANTHA") %>% unnest_tokens(word, dialogoue) %>% anti_join(stop_words) %>% count(word)

sam_dialogue%>% with(wordcloud(word,n))

sam_plot=sam_dialogue%>%arrange(desc(n))%>%filter(row_number()<10)%>% ggplot(aes(word,n,fill=word))+geom_col()+theme_pomological()+ggpomological::scale_fill_pomological()+theme(axis.title.x=element_blank(), axis.title.y=element_blank(),axis.text.x=element_blank(), axis.ticks.x=element_blank())


theo_dialogue<- her_script%>%filter(speaker=="THEODORE") %>% unnest_tokens(word, dialogoue) %>% anti_join(stop_words) %>% count(word)

theo_dialogue%>% with(wordcloud(word,n))

theo_plot=theo_dialogue%>%arrange(desc(n))%>%filter(row_number()<10)%>% ggplot(aes(word,n,fill=word))+geom_col()+theme_pomological()+ggpomological::scale_fill_pomological()+theme(axis.title.x=element_blank(), axis.title.y=element_blank(),axis.text.x=element_blank(), axis.ticks.x=element_blank())

sam_plot+theo_plot
```

```{r}
#set-up
most_scene<- her_script%>%group_by(scene,index) %>% unnest_tokens(word, dialogoue) %>% anti_join(stop_words) %>%ungroup()

#define fucntion:
plot_sentiment<-function(sent,col){
  nrc_emo <- get_sentiments("nrc") %>%  filter(sentiment == sent)
words <- most_scene %>%group_by(scene,index)%>% inner_join(nrc_emo) %>% count(word, sort = TRUE) %>%summarise(n=n())%>% arrange(index)
most_word=most_scene %>% inner_join(nrc_emo) %>% count(word, sort = TRUE) %>%arrange(desc(n))%>%filter(row_number()==1)
fin_plt=ggplot(words,aes(index,n,alpha=0.5))+geom_point(size=2,colour=col)+geom_line(colour=col)+ scale_fill_pomological() +geom_vline(aes(xintercept=mean(i)),alpha=0.4,colour="black",lty=4)+
  theme_pomological()+theme(legend.position = "none",axis.title.x=element_blank(), axis.title.y=element_blank(),axis.text.x=element_blank(), axis.ticks.x=element_blank())+geom_text(aes(label=most_word[[1]],x=max(index),y=max(n)),size=5)
return(fin_plt)

}
joy=plot_sentiment("joy","red")

sad=plot_sentiment("sadness","blue")

fear=plot_sentiment("fear","green")

anticipation=plot_sentiment("anticipation","purple")

pos=plot_sentiment("positive","red")

neg=plot_sentiment("negative","blue")
pos/neg

joy/sad

fear/anticipation
```
