---
title: "DIY Packages"
author: "Ashley Platt"
date: "May 10th 2022"
output: revealjs::revealjs_presentation
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r include=FALSE}
library(devtools)
library(roxygen2)
library(tidyverse)
```


## What do we need to make our own package? {data-background=#e5f6df}

- *Important libraries* 
    - roxygen2
    - devtools
- *A purpose*
    - e.g. a package which helps to streamline Tidy Tuesday analyses 

## What will our package contain? {data-background=#e5f6df}
- *functions*
  - help files
  - r files 
- *internal data*
- *analysis template*

## Let's make a package {data-background=#e5f6df}

- package is called *tthelp* 
- creating a package can be done;
  - via file --> new project --> new directory --> r package

- this creates a new directory which includes 
  - DESCRIPTION
  - NAMESPACE
  - two folders; man & R
  
- fun extra: add a start up message!
  
From here you can begin to build up your package. 

## Let's start with functions {data-background=#e5f6df}

- common purpose of creating packages is to hold your regularly used functions 
- I created 3 key functions for my package *tthelp*
  - *ttread*    :read in and summarise data from tidytuesdayR package
  - *ttwhich*   :indicate the most appropriate geoms based on x and y variable type
  - *ttgeom*    :plot a range of geoms
  
- Hint: use roxygensie() to create "help" page

  
**we can start with the basics of hello**

## Adding internal data {data-background=#e5f6df}

- another common purpose of user-generated packages is to self-contain your own data analysis
- packages can also hold internal data 
  - I used the *baby_names* TT data set to act as my internal data 
  
- internal data can be added by 
  - use_this :: use_data_raw("generate_internal")
  - this creates an R script in *data_raw* when you can put the code which     generates and save your own data


## Creating templates {data-background=#e5f6df}

- now I wanted to add a RMarkdown template within my package to help guide TT analysis and the use of my functions 

- used the code below to create a subdirectory within my package directory 
  - usethis::use_rmarkdown_template("Template Name")
  
## Creating templates cont. {data-background=#e5f6df}

- this creates 
  - *inst* folder 
    - *rmarkdown* folder
      - *templates* folder
        - *package_name* folder 
          - *skeleton* folder
            - *skeleton* rmd
- edit the rmd file to create your template 

- let's look at the template I made!

## Finished? {data-background=#e5f6df}

- use install() to install your package onto your computer
- now we can load this library into a new project, and use the templates and data it makes available 

e.g. devtools::install("/Users/ashleyplatt/Desktop/tthelp")


## Any questions? {data-background=#e5f6df}
  

