---
title: "Boardgames"
author: "Kate Riggall"
date: "03/02/2022"
output: 
  github_document:
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(magrittr)
library(tidyverse)
library(rvest)
library(lubridate)

mypal<-c("#3f3960", "#ff5100", "#ffffff", "#b4041e", "#3f84e5")
sliceValues <- rep(10, 5) 
pie(sliceValues,explode=0, theta=1.2, col=mypal)

```

#Prep
```{r load data, include=FALSE}
data<- tidytuesdayR::tt_load(2022, week = 4)
ratings <-data$ratings
details <-data$details

bgames<-merge(ratings,details, by="id")

bgames<-rename_with(bgames, ~ tolower(gsub(pattern="boardgame", replacement="", x=names(bgames))))


bgames[sapply(bgames, is.character)] <- lapply(bgames[sapply(bgames, is.character)], as.factor)
```


```{r extract entries}
bgames2<-bgames
bgames2<-bgames2%>%
  mutate(publisher_all = {str_replace_all(publisher, pattern = "\\[|\\]", replacement = "")}) %>%  
  separate(publisher_all, into = c("publisher"), sep = ",") %>% 
  mutate(publisher = gsub("[[:punct:]]", "", publisher))

bgames2<-bgames2 %>%
  mutate(artist_all = {str_replace_all(artist, pattern = "\\[|\\]", replacement = "")}) %>%  
  separate(artist_all, into = c("artist"), sep = ",") %>% 
  mutate(artist = gsub("[[:punct:]]", "", artist))

bgames2<-bgames2 %>%
  mutate(category_all = {str_replace_all(category, pattern = "\\[|\\]", replacement = "")}) %>%  
  separate(category_all, into = c("category"), sep = ",") %>% 
  mutate(category = gsub("[[:punct:]]", "", category))

bgames2<-bgames2 %>%
  mutate(category_all = {str_replace_all(category, pattern = "\\[|\\]", replacement = "")}) %>%  
  separate(category_all, into = c("category"), sep = ",") %>% 
  mutate(category = gsub("[[:punct:]]", "", category))

bgames2<-bgames2 %>%
  mutate(mechanic_all = {str_replace_all(mechanic, pattern = "\\[|\\]", replacement = "")}) %>%  
  separate(mechanic_all, into = c("mechanic"), sep = ",") %>% 
  mutate(mechanic = gsub("[[:punct:]]", "", mechanic))

variables<-names(bgames2)
```


```{r new variables}
bgames2$timebracket<-cut(bgames2$playingtime, breaks = c(0,60,120,180,240,300,360,420,480,540 ), include.lowest=TRUE, ordered_result = TRUE, labels = c("<60", "<120", "<180", "<240", "<300", "<360", "<420", "<480", "<540"))

bgames2$mintimebracket<-cut(bgames2$minplaytime, breaks = c(0,60,120,180,240,300,360,420,480,540 ), include.lowest=TRUE, ordered_result = TRUE, labels = c("<60", "<120", "<180", "<240", "<300", "<360", "<420", "<480", "<540"))

summary(bgames2$timebracket)
summary(bgames2$mintimebracket)

```


```{r plot_predictors}
#plot average rating by playing time and minimum number of players
ggplot(bgames2, aes(playingtime, minplayers)) +
  geom_tile(aes(fill=average),na.rm=TRUE) +
  scale_fill_gradient2(low=mypal[4], mid=mypal[3],high=mypal[5], midpoint=5)
```

```{r plot_predictors2}

#turns out there are some games you can play forever? may be a data quality issue - for now let's cut that variable (I do this up the top so it goes into all dataframes)

ggplot(bgames2, aes(timebracket, minplayers)) +
  geom_tile(aes(fill=average),na.rm=TRUE) +
  scale_fill_gradient2(low=mypal[4], mid=mypal[3],high=mypal[5], midpoint=5)
```

```{r plot_predictors3}
#why is this not working?
hist(bgames2$average)
```

```{r plot_predictors4}
#data is skewed!

#is bayes average better?
hist(bgames2$bayes_average)
```

```{r plot_predictors_standardise}
#no, that's worse.

#standardize the original average (just for this plot)
bgames2$st_average<- (bgames2$average - mean(bgames2$average)) / sd(bgames2$average)

#plot the standardised average
ggplot(bgames2, aes(timebracket, minplayers)) +
  geom_tile(aes(fill=st_average),na.rm=TRUE) +
  scale_fill_gradient2(low=mypal[4], mid=mypal[3],high=mypal[5], midpoint=0)
```


```{r plot_predictors_no_outliers}
#or we could just eliminate outliers, and set the midpoint to the median for the data
boxplot(bgames2$average)

Q <- quantile(bgames2$average, probs=c(.25, .75), na.rm = FALSE)

iqr <- IQR(bgames2$average)

up <-  Q[2]+1.5*iqr # Upper Range  
low<- Q[1]-1.5*iqr # Lower Range

eliminated<- subset(bgames2, bgames2$average > (Q[1] - 1.5*iqr) & bgames2$average< (Q[2]+1.5*iqr))

g<-ggplot(eliminated, aes(timebracket, minplayers)) +
  geom_tile(aes(fill=average),na.rm=TRUE) +
  scale_fill_gradient2(low=mypal[4], mid=mypal[3],high=mypal[5], midpoint=median(eliminated$average))+
  labs(title="Average User Ratings by Minimum Number of Players and Estimated Play Time", x="Estimated Play Time (Bracketed)", y="Minimum Players")
g
```

```{r plot_predictors_fix_scale}


#this is okay, but not really making use of the colour scale - we can adjust the colour scale to improve (thanks Hayley for reminding me!)

g<-ggplot(eliminated, aes(timebracket, minplayers)) +
  geom_tile(aes(fill=average),na.rm=TRUE) +
  scale_fill_gradient2(low=mypal[4], mid=mypal[3],high=mypal[5], midpoint=7, limits=c(5,9))+
  labs(title="Average User Ratings by Minimum Number of Players and Estimated Play Time", x="Estimated Play Time (Bracketed)", y="Minimum Players")
g
```

```{r plot_predictors_dark}
#finally, since there are gaps in the plot, adjust the theme so the data is more obvious

g<-g +
  theme(panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),panel.background = element_rect(fill="grey30"))+
  labs(title="Average User Ratings by Minimum Number of Players and Estimated Play Time", x="Estimated Play Time (Bracketed)", y="Minimum Players")
g

```

```{r plot_predictors_tabletop}
#just for fun let's make it look like a tabletop

library(ggpubr)
library(jpeg)
# Import the image
img <- readJPEG("background-texture.jpeg")
  
# Plot with background image [must go first]
g<-ggplot(eliminated, aes(timebracket, minplayers)) +
  background_image(img)+
  geom_tile(aes(fill=average),na.rm=TRUE) +
  scale_fill_gradient2(low=mypal[4], mid=mypal[3],high=mypal[5], midpoint=7, limits=c(5,9))+
  labs(title="Average User Ratings by Minimum Number of Players and Estimated Play Time", x="Estimated Play Time (Bracketed)", y="Minimum Players")
g
```

This is interesting, but keep in mind that these ratings are given by players who are very interested in boardgames generally, who have already decided to buy or play the game, and then bothered to rate it. Let's look at how many users rated each game.
```{r users_rated}
g<-ggplot(bgames2, aes(timebracket, minplayers)) +
  background_image(img)+
  geom_tile(aes(fill=users_rated),na.rm=TRUE) +
  scale_fill_gradient2(low=mypal[4], mid=mypal[3],high=mypal[5], midpoint=median(bgames2$users_rated))+
labs(title="Number of Users Who Rated",subtitle = "by minimum players and typical play time")
g
```

```{r users_rated2}
hist(bgames2$users_rated)
```

```{r users_rated_no_outliers}
Q <- quantile(bgames2$users_rated, probs=c(.25, .75), na.rm = FALSE)

iqr <- IQR(bgames2$users_rated)

up <-  Q[2]+1.5*iqr # Upper Range  
low<- Q[1]-1.5*iqr # Lower Range

eliminated<- subset(bgames2, bgames2$users_rated > (Q[1] - 1.5*iqr) & bgames2$users_rated< (Q[2]+1.5*iqr))

hist(eliminated$users_rated)
```

```{r users_rated_no_outliers2}
hist(log(eliminated$users_rated))
```

```{r users_rated_no_outliers3}
g<-ggplot(eliminated, aes(timebracket, minplayers)) +
  background_image(img)+
  geom_tile(aes(fill=log(users_rated)),na.rm=TRUE) +
  scale_fill_gradient2(low=mypal[1], mid=mypal[3],high=mypal[2], midpoint=5,  limits=c(3,7))+
labs(title="Number of Users Who Rated",subtitle = "by minimum players and typical play time")
g

```
it would be good to look at some categorical variables, but there are too many categories in each, so let's take the top 10

```{r top_10_lists}
top10<-function(data, variable) {
  top10list<-sort(table(data[[variable]]),decreasing=TRUE)[1:10]
  names(top10list)
  }
  
top10pub<-top10(bgames2,"publisher")
top10cat<-top10(bgames2, "category")
top10mech<-top10(bgames2, "mechanic")
top10art<-top10(bgames2, "artist")

```


#Publishers by country

```{r publisher_countries}
url <-  "https://boardgamegeek.com/wiki/page/Game_Manufacturers"

df <- url %>% 
  read_html() %>% 
  html_nodes("table") %>% 
  html_table(fill = T) %>% 
  lapply(., function(x) setNames(x,c("publisher", "country", "website"))) 

pub_countries <- df[[2]]

pub_countries<-inner_join(bgames2,pub_countries, by="publisher") 

pub_countries$country<-as.factor(pub_countries$country)
pub_countries$country<-droplevels(pub_countries$country)
levels<-levels(pub_countries$country) [1:6]
levels(pub_countries$country)<-c(levels,levels[2],"International")

summary(pub_countries$country)
```

```{r publisher_countries2}

ggplot(pub_countries, aes(country)) +
  geom_bar()+
  labs(title="Total Games Published By Country", subtitle="By Currently Active Publishers", x="Country", y="Games")
```
let's try breaking this down by category
```{r publisher_countries_category}
ggplot(pub_countries[pub_countries$category %in% top10cat,], aes(country, category))+geom_count()
```
I would like to try to use some fun shapes - unicode should work but doesn't (see code commented out), so for now I will use the standard shapes
```{r publisher_countries_category2}
#doesn't work
# ggplot(pub_countries[pub_countries$category %in% top10cat,], aes(country, category, shape=factor(category))) +
#   geom_count()+
#   scale_shape_manual(values = c("\u1F9E0", "\u1F590", "\u26F0", "\u1F3DB", "\u1F43E", "\u1F610", "\u1F0A0", "\u1F3B2", "\u1F4B0", "\u2694"))

ggplot(pub_countries[pub_countries$category %in% top10cat,], aes(country, category, shape=factor(category))) +
  geom_count()+
  scale_shape_manual(values=c(1:10))
```
Okay then, just for fun let's colour it like a Go board - black and white (by popularity)
```{r publisher_countries_category3}
category_ratings<-pub_countries %>%
  group_by(category) %>%
  summarise(mean_rating = mean(average))

pub_countries<-left_join(pub_countries, category_ratings)

pub_countries<-pub_countries %>%
  mutate(popular=ifelse(mean_rating>mean(mean_rating),"Yes", "No"))

ggplot(pub_countries[pub_countries$category %in% top10cat,], aes(country, category, colour=popular))+ geom_count() +
  scale_colour_manual(values=c("black","white")) +labs(title="Common and Popular Categories by Country")+
  theme(axis.text.x=element_text(angle =90, vjust=.5))
```


#Publishing Trends By Year/Country

We only have one timepoint for the boardgame ratings, but because we have publishing year, we can look at what kinds of games are being published using a frequency plot (adapted from a plot of word use frequency on different twitter feeds)

I initially thought there was a data quality problem with the publishing dates

```{r}
boxplot(bgames$yearpublished)
```
but then I remembered that a few boardgames were invented in Ancient Egypt and Sumeria, and yep, turns out:

```{r}
subset(bgames2,yearpublished<(-2000), select=name)
```

So I have restricted it to games published after 1980; I'm also using the dataframe with countries so that I can look at one at a time
```{r minplayers}
pub_countries$minplayers<-as.factor(pub_countries$minplayers)
minplay_by_time <- subset(pub_countries,yearpublished>1980) %>%
  count(yearpublished, country, minplayers) %>%
  group_by(country, yearpublished) %>%
  mutate(time_total = sum(n)) %>%
  group_by(minplayers) %>%
  mutate(minplayers_total = sum(n)) %>%
  ungroup() %>%
  rename(count = n) %>%
  filter(minplayers_total > 30)

minplay_by_time 
```


```{r}
nested_data <- minplay_by_time %>%
  nest(data = c(-minplayers, -country))

nested_data
```


```{r}

library(purrr)

nested_models <- nested_data %>%
  mutate(models = map(data, ~ glm(cbind(count, time_total) ~ yearpublished, ., 
                                  family = "binomial")))

```


```{r}
library(broom)

slopes <- nested_models %>%
  mutate(models = map(models, tidy)) %>%
  unnest(cols = c(models)) %>%
  filter(term == "yearpublished") %>%
  mutate(adjusted.p.value = p.adjust(p.value))

slopes
```




```{r}
minplay_by_time %>%
  inner_join(slopes, by = c("minplayers", "country")) %>%
  filter(country == "United States") %>%
  ggplot(aes(yearpublished, count/time_total, color = minplayers)) +
  geom_line(size = 1.3) +
  labs(x = NULL, y = "Frequency") +
  labs(title="Publishing Trends over Time",subtitle="Minimum Player Number", x="Year")+
  scale_colour_discrete(name="Minimum Players")

```


```{r}
minplay_by_time <- subset(pub_countries,yearpublished>1980) %>%
  count(yearpublished, country, publisher) %>%
  group_by(country, yearpublished) %>%
  mutate(time_total = sum(n)) %>%
  group_by(publisher) %>%
  mutate(publisher_total = sum(n)) %>%
  ungroup() %>%
  rename(count = n) %>%
  filter(publisher_total > 30)

minplay_by_time 
```


```{r}
nested_data <- minplay_by_time %>%
  nest(data = c(-publisher, -country))

nested_data
```


```{r}

library(purrr)

nested_models <- nested_data %>%
  mutate(models = map(data, ~ glm(cbind(count, time_total) ~ yearpublished, ., 
                                  family = "binomial")))

```


```{r}
library(broom)

slopes <- nested_models %>%
  mutate(models = map(models, tidy)) %>%
  unnest(cols = c(models)) %>%
  filter(term == "yearpublished") %>%
  mutate(adjusted.p.value = p.adjust(p.value))

slopes
```

```{r}
minplay_by_time %>%
  inner_join(slopes, by = c("publisher", "country")) %>%
  filter(country == "United States") %>%
  ggplot(aes(yearpublished, count/time_total, color = publisher)) +
  geom_line(size = 1.3) +
  labs(x = NULL, y = "Frequency") +
  labs(title="Publisher Output over Time",subtitle="United States", x="Year")+
  scale_colour_discrete(name="Publisher")

```

```{r}
minplay_by_time %>%
  inner_join(slopes, by = c("publisher", "country")) %>%
  filter(country == "Germany") %>%
  ggplot(aes(yearpublished, count/time_total, color = publisher)) +
  geom_line(size = 1.3) +
  labs(x = NULL, y = "Frequency")+
  labs(title="Publisher Output over Time",subtitle="Germany", x="Year")+
  scale_colour_discrete(name="Publisher")

```
