---
title: "2022-12-06_open-psychometrics_HC"
author: "Hayley Caldwell"
date: "2022-11-29"
output: github_document
---

```{r setup, include=FALSE}

pacman::p_load(install = FALSE, update = FALSE, tidytuesdayR, tidyverse, tidyEmoji, dplyr, ggimage, scales, wordcloud, fmsb)

tt_data_op <- tt_load("2022-08-16")

characters <- tt_data_op$characters
psych_stats <- tt_data_op$psych_stats
myers_briggs <- tt_data_op$myers_briggs

setwd("E:/Tidy_Tuesday")

```

The Open Psychometrics project involves a website which provides a collection of interactive personality tests, one of which is the Statistical "Which Character" Personality Quiz. This quiz takes your subjective ratings from a selection of dichotomous traits, then correlates them with the ratings given to various TV and film characters, to then list which characters your personality scores correlate with. The ratings of characters have been taken from ratings given by members of the public on the website. 800 of these characters from a randomly selected 100 universes are included in these data. 

```{r op data}

# view(characters)
# view(psych_stats) 

```


```{r clean}

# unique(psych_stats$question) 
  # 400 questions, including 35 emoji questions

# get rid of emojis 
psych_stats1 <- psych_stats %>% 
  filter(grepl("[A-Za-z]", question))

characters1 <- characters %>% 
  rename("char_name" = name) %>% 
  rename("char_id" = id) %>% 
  select(-uni_name, -char_id, -uni_id) %>% 
  group_by(char_name) %>% 
  slice_max(notability) %>%  # multiple entries per char for different shows 
  ungroup()

```


```{r ratings}

# which characters and shows have been rated the most?

# combine charaters and psych_stats 
char_psych <- left_join(characters, psych_stats1)

# rank shows by number of votes, then cut off first and last 15
show_ratings <- char_psych %>% 
  group_by(uni_name) %>% 
  mutate(mean_votes = mean(number_ratings)) %>% 
  select(uni_name, mean_votes) %>% 
  distinct() %>% 
  arrange(-mean_votes)

# top 15
shows_t15 <- show_ratings %>% 
  head(15)

# bottom 15
shows_b15 <- show_ratings %>% 
  tail(15)

shows_bt15 <- full_join(shows_b15, shows_t15)

# bar graph number of votes per show 
show_ratings_bt15.g <- ggplot(shows_bt15, aes(x = reorder(uni_name, mean_votes), y = mean_votes, fill = uni_name)) +
  geom_col() +
  coord_flip() +
  theme_bw() +
  theme(legend.position = "none") +
  xlab("Universe")+
  ylab("Mean Votes") +
  annotate("text", y = 950, x = 23, label = "Top 15") +
  annotate("text", y = 970, x = 8, label = "Bottom 15") +
  geom_image(mapping = aes(x = 23, y = 800, image = 'bracket.png'), size = 0.13) +
  geom_image(mapping = aes(x = 8, y = 800, image = 'bracket.png'), size = 0.13)  + 
  scale_y_continuous(limits=c(0,1100))
show_ratings_bt15.g

```

```{r notability}

# which show has most notable characters?

characters %>% 
  arrange(-notability) %>% 
  head(20)

notable_shows <- characters %>% 
  group_by(uni_name) %>% 
  mutate(mean_notability = mean(notability)) %>% 
  select (uni_name, mean_notability) %>% 
  distinct() %>% 
  arrange(-mean_notability)

# top 15
notable_shows_t15 <- notable_shows %>% 
  head(15)

# bottom 15
notable_shows_b15 <- notable_shows %>% 
  tail(15)

notable_shows_bt15 <- full_join(notable_shows_b15, notable_shows_t15)

# bar graph notability for top and bottom shows 
notable_shows_bt15.g <- ggplot(notable_shows_bt15, aes(x = reorder(uni_name, mean_notability), y = mean_notability, fill = uni_name)) +
  geom_col() +
  coord_flip() +
  theme_bw() +
  theme(legend.position = "none") +
  xlab("Universe") +
  ylab("Mean Notability") +
  annotate("text", y = 105, x = 23, label = "Top 15") +
  annotate("text", y = 108, x = 8, label = "Bottom 15") +
  geom_image(mapping = aes(x = 23, y = 85, image = 'bracket.png'), size = 0.13) +
  geom_image(mapping = aes(x = 8, y = 85, image = 'bracket.png'), size = 0.13) + 
  scale_y_continuous(limits=c(0,120))
notable_shows_bt15.g
```

```{r notability relationship with mean votes}

# notability and mean votes relationship?

notable_show_ratings <- left_join(notable_shows, show_ratings)

notable_show_ratings.g <- ggplot(notable_show_ratings, aes(x = mean_notability, y = mean_votes)) +
  geom_point(colour = "#EB941F") +
  geom_smooth(colour = "#10BE48") +
  theme_bw() +
  xlab("Notability") +
  ylab("Votes") +
  theme(legend.position = "none")
notable_show_ratings.g

```


```{r highest rated characters for traits}

# which character has the highest rating for ~30 OP traits

# choose ~30 fun traits 
# unique(psych_stats1$personality)

fun_stats <- c("neat", "messy", "practical", "imaginative", "dominant", "submissive", "beautiful", "ugly", "loud", "quiet", "psychopath", "empath", "high IQ", "low IQ", "nerd", "jock", "feminine", "masculine", "weird", "normal", "chaotic", "orderly", "optimistic", "pessimistic", "Coke", "Pepsi", "introvert", "extrovert", "stinky", "fresh", "'right-brained'", "'left-brained'", "German", "English")

fun_stats1 <- c("messy", "imaginative", "dominant", "beautiful", "loud", "empath", "high IQ", "jock", "feminine", "normal", "chaotic", "optimistic", "Coke", "extrovert", "fresh", "'left-brained'", "English")
fun_stats2 <- c("neat", "practical", "submissive", "ugly", "quiet", "psychopath", "low IQ", "nerd", "masculine", "weird", "orderly", "pessimistic", "Pepsi", "introvert", "stinky", "'right-brained'", "German")

```

``` {r highest rated characters for traits plot}

# group by trait, slice max, plot photos and names on graph
psych_stats_34 <- psych_stats1 %>% 
  filter(personality %in% fun_stats) %>% 
  group_by(personality) %>% 
  slice_max(avg_rating) %>% 
  mutate(avg_rating_di = case_when(
    personality %in% fun_stats1 ~ (avg_rating *-1),
    personality %in% fun_stats2 ~ avg_rating),
    group = case_when(
    personality %in% fun_stats1 ~ "1",
    personality %in% fun_stats2 ~ "2"),
    avg_rating_label = case_when(
      avg_rating_di <0 ~ (avg_rating_di + 40),
      avg_rating_di >0 ~ (avg_rating_di - 40)
    )) %>% 
  filter(!char_name %in% c("Alexis Rose", "Sam Obisanya", "Ilana Wexler")) %>% # ties
  ungroup()

psych_char_34 <- left_join(psych_stats_34, characters1)
  
# plot side by side 
psych_char_34.g <- ggplot(psych_char_34, aes(x = question, y = avg_rating_di, fill = group)) +
  geom_col() +
  theme_bw() +
  theme(legend.position = "none") +
  coord_flip() +
  xlab("Question") +
  ylab("Rating") +
  geom_hline(aes(yintercept=0)) +
  scale_fill_manual(values=c("#EB941F", "#10BE48")) +
  geom_image(mapping = aes(image = image_link)) +
  geom_text(aes(y = avg_rating_label, x = question, label = char_name))
psych_char_34.g

```

```{r word coud hamilton, echo=TRUE, warning=FALSE}

# word clouds to compare very different TV shows on what psychometric traits define them

# pick 3 very different TV shows
# unique(characters$uni_name)

# hamilton
hamilton_ratings <- char_psych %>% 
  filter(uni_name %in% c("Hamilton")) %>% 
  group_by(personality) %>% 
  mutate(mean_rating = mean(avg_rating)) %>% 
  select(uni_name, mean_rating, personality) %>% 
  distinct()%>% 
  with(wordcloud(personality, mean_rating, random.order=FALSE, scale=c(2,0.1), colors=brewer.pal(100, "Dark2")))

```

```{r word cloud harry potter, echo=TRUE, warning=FALSE}

# harry potter 
hp_ratings <- char_psych %>% 
  filter(uni_name %in% c("Harry Potter")) %>% 
  group_by(personality) %>% 
  mutate(mean_rating = mean(avg_rating)) %>% 
  select(uni_name, mean_rating, personality) %>% 
  distinct() %>% 
  with(wordcloud(personality, mean_rating, random.order=FALSE, scale=c(2,0.1), colors=brewer.pal(100, "YlOrBr")))

```

```{r word cloud star wars, echo=TRUE, warning=FALSE}

#star wars 
starwars_ratings <- char_psych %>% 
  filter(uni_name %in% c("Star Wars")) %>% 
  group_by(personality) %>% 
  mutate(mean_rating = mean(avg_rating)) %>% 
  select(uni_name, mean_rating, personality) %>% 
  distinct() %>% 
  with(wordcloud(personality, mean_rating, random.order=FALSE, scale=c(2,0.1), colors=brewer.pal(100, "Spectral")))

```

```{r spider web spongebob}

# spider web graphs

# spongebob vs squidward

sb_stats_10 <- c("messy/neat", "imaginative/practical", "beautiful/ugly", "loud/quiet", "high IQ/low IQ", "optimistic/pessimistic", "extrovert/introvert", "fresh/stinky", "comedic/dramatic", "cruel/kind")

sb_stats <- psych_stats1 %>% 
  filter(question %in% sb_stats_10,
         char_name %in% c("SpongeBob SquarePants", "Squidward Tentacles")) %>%
  mutate(scaled_rating = case_when(
    personality %in% c("introvert", "neat", "practical", "ugly", "quiet", "low IQ", "pessimistic", "stinky", "dramatic", "kind") ~ (100 - avg_rating),
    personality %in% c("extrovert", "messy", "imaginative", "beautiful", "loud", "high IQ", "optimistic", "fresh", "comedic", "cruel") ~ avg_rating
  )) %>% 
  select(char_name, question, scaled_rating) %>% 
  pivot_wider(names_from = question, values_from = scaled_rating) %>% 
  select(!char_name)

# plot
rownames(sb_stats) <- c("SpongeBob", "Squidward") #Spongebob = orange, Squidward = green

sb_stats <- rbind(rep(100,25) , rep(0,25) , sb_stats)

coul <- c("#EB941F", "#10BE48")
colours_border <- coul
colours_in <- alpha(coul,0.3)

radarchart(sb_stats, axistype=1, plwd=4, plty=1, cglty=1, cglwd=0.8, vlcex=0.8,
    pcol=colours_border, pfcol=colours_in, 
    cglcol="grey", axislabcol="grey", caxislabels=seq(0, 100, 25))

# legend(x=1, y=1, legend = rownames(sb_stats[-c(1,2),]), bty = "n", pch=20 , col=colours_in , text.col = "grey", cex=1.2, pt.cex=3)

```

```{r spider web star wars}

# Darth Vader vs Luke Skywalker 

sw_stats <- psych_stats1 %>% 
  filter(question %in% sb_stats_10,
         char_name %in% c("Luke Skywalker", "Darth Vader")) %>%
  mutate(scaled_rating = case_when(
    personality %in% c("introvert", "neat", "practical", "ugly", "quiet", "low IQ", "pessimistic", "stinky", "dramatic", "kind") ~ (100 - avg_rating),
    personality %in% c("extrovert", "messy", "imaginative", "beautiful", "loud", "high IQ", "optimistic", "fresh", "comedic", "cruel") ~ avg_rating
  )) %>% 
  select(char_name, question, scaled_rating) %>% 
  pivot_wider(names_from = question, values_from = scaled_rating) %>% 
  select(!char_name)

# plot
rownames(sw_stats) <- c("Vader", "Luke") # vader = red, Luke = blue

sw_stats <- rbind(rep(100,25) , rep(-0,25) , sw_stats)

coul2 <- c("#FB4A33", "#3398FB")
colours_border2 <- coul2
colours_in2 <- alpha(coul2,0.3)

radarchart(sw_stats, axistype=1, plwd=4, plty=1, cglty=1, cglwd=0.8, vlcex=0.8,
    pcol=colours_border2, pfcol=colours_in2, 
    cglcol="grey", axislabcol="grey", caxislabels=seq(0, 100, 25))

# legend(x=1, y=1, legend = rownames(sw_stats[-c(1,2),]), bty = "n", pch=20 , col=colours_in2 , text.col = "grey", cex=1.2, pt.cex=3)

```


There was also data about the same character's in relation to their Myers-Briggs scores. Myers Briggs is a common personality measurement which provides takers with a 4 letter personality code based on 4 dichotomous trait pairs (Extrovert-Introvert, Sensing-Intuiting, Thinking-Feeling, and Judging-Perceiving). The data includes the average percentage match of test-takers in the open-psychometrics quiz, with each Myers-Briggs personality type which test-takers self-identified after the quiz.

```{r myers brigs data}

# view(myers_briggs)

```


```{r myers briggs cleaning}

myers_briggs_top <- myers_briggs %>% 
  group_by(char_name) %>% 
  slice_max(avg_match_perc) %>%
  slice_max(number_users) %>% # multiple entries per char for different shows
  mutate(myers_briggs_total = myers_briggs) %>% 
  separate(myers_briggs, into = c("0", "Extrovert_Introvert", "Sensing_Intuiting", "Thinking_Feeling", "Judging_Perceiving"), sep = "") %>% 
  select(-"0") %>% 
  ungroup() %>% 
  distinct() 

```

```{r myers briggs overall type count}

# how many of each overall type 

myers_briggs_all <- myers_briggs_top %>% 
  group_by(myers_briggs_total) %>% 
  mutate(myers_briggs_total_count = n()) %>% 
  select(myers_briggs_total, myers_briggs_total_count) %>% 
  distinct()

myers_briggs_all.g <- ggplot(myers_briggs_all, aes(x = reorder(myers_briggs_total, myers_briggs_total_count) , y = myers_briggs_total_count, fill = myers_briggs_total)) +
  geom_col() +
  theme_bw() +
  theme(legend.position = "none") +
  coord_flip() +
  xlab("Myers Briggs Personality") +
  ylab("Frequency")
myers_briggs_all.g

```

```{r myers briggs individual result count}

# how many characters of each trait are there

myers_briggs_count1 <- myers_briggs_top %>% 
  pivot_longer(cols = c(Extrovert_Introvert, Sensing_Intuiting, Thinking_Feeling, Judging_Perceiving), names_to = 'scale', values_to = "result") %>% 
  group_by(result) %>% 
  mutate(result_freq = n()) %>%
  ungroup() 

myers_briggs_count1$result_freq <- as.double(myers_briggs_count1$result_freq)

myers_briggs_count <- myers_briggs_count1 %>% 
  mutate(result_freq_scaled = case_when(
    result %in% c("E", "S", "T", "J") ~ (result_freq*-1),
    result %in% c("I", "N", "F", "P") ~ result_freq),
    group = case_when(
    result %in% c("E", "S", "T", "J") ~ 1,
    result %in% c("I", "N", "F", "P") ~ 2)  
    ) %>% 
  select(scale, result, result_freq_scaled, group) %>% 
  distinct()
  
myers_briggs_count.g <- ggplot(myers_briggs_count, aes(x = scale, y = result_freq_scaled, fill = group)) +
  geom_col() +
  theme_bw() +
  theme(legend.position = "none") +
  coord_flip() +
  xlab("Question") +
  ylab("Frequency") +
  geom_hline(aes(yintercept=0)) 
myers_briggs_count.g

```

See mb_app.R for the Myers-Briggs Shiny app
